---
title: "Analysis of Breast Cancer data using ggplots"
author: "Praveen Nagarajan"
date: "4/28/2019"
output: 
  html_document:
    toc: true
---

```{r echo=FALSE}
library("ggplot2")

data_set <- read.csv("data.csv")

colnames(data_set)[colnames(data_set)=="diagnosis"] <- "Diagnosis"
data_set$Diagnosis <- factor(data_set$Diagnosis, labels=c("Benign", "Malignant"))
```

## 1. geom_point

### i. Initial Visualization
``` {r graph1-geom_point}
ggplot(data_set, aes(x = perimeter_mean, y = concavity_mean)) +geom_point(colour="black") +labs(x = "Perimeter", y = "Concavity")
```
<p>The first graph above shows a strong positive relationship between the Perimeter of a particular tumor sample and its Concavity.</p>
<p>The Perimeter of a particular tumor sample is the mean size of the tumor's core.</p>
<p>The Concavity expresses the mean severity of the concave portions of the tumor's contour.</p>
<p>We will be exploring this relationship further using the variations of the ggplot attributes.</p>

### ii. Indicate points as Benign/Malignant
```{r graph2-geom_point_colour-Diagnosis, echo=FALSE}
ggplot(data_set, aes(x = perimeter_mean, y = concavity_mean, colour=Diagnosis)) +geom_point() +labs(x = "Perimeter", y = "Concavity")
```
<p>In this above graph, we add different colors for the Diagnosis levels (Benign, Malignant).</p>
<p>We can visually see a relationship between Perimeter, Concavity, and the Diagnosis.</p>
<p>Note that Benign is colored Red, and Malignant is colored Teal.</p>

### iii. Seperate the points labeled Benign/Malignant
``` {r graph3-geom_point_facet_wrap-Diagosis, echo=FALSE}
ggplot(data_set, aes(x = perimeter_mean, y = concavity_mean, colour=Diagnosis)) +geom_point() +facet_wrap(data_set$Diagnosis) +labs(x = "Perimeter", y = "Concavity")
```
<p>This above graph, shows the tumor samples seperated by their classification as Benign/Malignant.</p>
<ol>We can see a few trends:
<li> Benign examples are clustered with low perimeter and concavity.</li>
<li> Malignant examples have high perimeter (which feels intuitive).</li>
<li>
<ul>Numerical observations: 
<li> Benign cases have Perimeter less than 120.</li>
<li> Malignant cases have Perimeter greater than 60.</li>
</ul>
</li>
</ol>

## 2. geom_smooth

### i. loess - "Locally Estimated Scatterplot Smoothing" visualization
``` {r graph4-geom_smooth}
ggplot(data_set, aes(x = perimeter_mean, y = concavity_mean)) +geom_point() +geom_smooth(span=.5) +labs(x = "Perimeter", y = "Concavity")
```
<p>This plot overlays a loess-fitted curve to the scatterplot.</p>
<p>We can see the fitted line and the estimated confidence.</p>

### ii. loess visualization seperated by Benign/Malignant
``` {r graph5-geom_smooth_colour-Diagnosis}
ggplot(data_set, aes(x = perimeter_mean, y = concavity_mean, colour=Diagnosis)) +geom_smooth(span=.5) +labs(x = "Perimeter", y = "Concavity")
```
<p>We can see a fitted line generated for both Benign cases and Malignant cases.</p>
<p>It is interesting to note that with a span of just 0.5 the loess curves for Benign and Malignant (including the areas for confidence) may be linearly seperable. </p>

## 3. geom_histogram

### i. Histogram of Perimeter with Benign/Malignant classifications
``` {r graph6-geom_histogram_colour-Diagnosis}
ggplot(data_set, aes(perimeter_mean, colour=Diagnosis)) +geom_histogram(binwidth=1) +labs(x="Perimeter")
```

### ii. Seperate Histograms of Perimeter with Benign/Malignant classifications
``` {r graph7-geom_histogram_facet_wrap-Diagnosis}
ggplot(data_set, aes(perimeter_mean, colour=Diagnosis)) +geom_histogram(binwidth=1) +facet_wrap(data_set$Diagnosis) +labs(x="Perimeter")
```
